*Prochlorococcus marinus* responses to light and oxygen

### 2024-06-10 Code/Manuscript_O2.Rmd
Knit with new map

Similar to growth rate trends under 250 µM O~2~, MED4 maintained at 25 µM O~2~ showed fastest growth rate when the photoperiod was 16 h for each spectral waveband, across PAR levels (Fig \@ref(fig:MED4PARGAM)). The
µ~max~ was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 180 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod experiments under white LED light were not performed as no growth was achieved when grown under an 8 h photoperiod of white LED light.

@@

The growth rates for *Prochlorococcus marinus* SS120 clade LLII/III, under 250 µM O~2~, 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands,
increased with longer photoperiods (Fig \@ref(fig:SS120PARGAM)).
No growth was observed under any blue light photoperiods when exposed to
peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however
increased with increasing photoperiods for white and red light under
peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at
16 h red light photoperiod. Growth rate decreased with longer
photoperiods and showed growth inhibition at 12 and 16 h photoperiods
under PAR of 180 µmol photons m^-2^ s^-1^ white LED, red or blue light.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^ growth rate trends
were similar to 250 µM O~2~. SS120 showed no growth under a 4 h
photoperiod for red spectral waveband, however under blue light, SS120
was able to grow (Fig \@ref(fig:SS120PARGAM)). In contrast to the growth rate
trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^
experiments, SS120 grew under 4 and 8 h blue light and 16 h red light
photoperiods, however the growth rate decreased under 12 and 16 h white
LED light photoperiod treatments. Blue light treatments under PAR of 180
µmol photons m^-2^ s^-1^ showed growth only under an 8 h photoperiod.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16
h photoperiod and 180 µmol photons m^-2^ s^-1^ under white LED light
experiments were not performed due to time constraints.

SS120 did not reproducibly grow when sparged to the lowest O~2~ of 2.5
µM (Fig \@ref(fig:SS120PARGAM)). 2.5 µM O~2~ growth rate experiments were not
conducted for 4 and 16 h photoperiods under peak PAR of 180 µmol photons
m^-2^ s^-1^ and for red light under peak PAR of 90 µmol photons
m^-2^ s^-1^, as no growth occurred when SS120 was exposed to 8 and 12 h
photoperiods.

@@

#  This is Fig10.tiff
NichesMap_cap <- glue("**Potential Future Niches for *Prochlorococcus* clades in a warming ocean.**",
                      "The maps shows current latitudinal distribution limits for *Prochlorococcus* (between green dashed lines), along with tropical and temperate  latitudinal regions (yellow bands).",
"Light levels, photoperiods and spectral bands, potentially permissive for growth of *Prochlorococcus* clades are determined by the interactions of latitude, season and depth attenuation of light.",
"**A.** Potential clade occupancies under Full Oxygen (250 µM O~2~) are shown for 10-50 m and 100 m depths, in temperate vs. tropical regions, under depth-resolved photoperiods, approximating Winter Solstice, Equinox and Summer Solstice. Clade HLI is excluded from temperate depths by short photoperiods at Winter Solstice. Clades LLII/III and LLIV are excluded from shallow depths by excessive light.",
"**B.** Low Oxygen Zones (25 µM O~2~) drastically alter these potential occupancy patterns by excluding clade HLI from the temperate winter solstice at all depths, and by expanding the exploitation of higher light zones by Clade LLIV.",
"Potential clade occupancies for the southern temperate zone mirror the northern temperate zone occupancies.",

>>>>

Similar to growth rate trends under 250 µM O~2~, MED4 maintained at 25 µM O~2~ showed fastest growth rate when the photoperiod was 16 h for each spectral waveband, across PAR levels (Fig \@ref(fig:MED4PARGAM)). The
µ~max~ was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 12) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ blue light and 12 h photoperiod. The 4 h photoperiod experiments under white LED light were not performed as no growth was achieved when grown under an 8 h photoperiod of white LED light.

@@

The growth rates for *Prochlorococcus marinus* SS120 clade LLII/III, under 250 µM O~2~, 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands,
increased with longer photoperiods (Fig \@ref(fig:SS120PARGAM)A).
No growth was observed under any blue light photoperiods when exposed to
peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however
increased with increasing photoperiods for white and red light under
peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at
16 h red light photoperiod. Growth rate decreased with longer
photoperiods and showed growth inhibition at 12 and 16 h photoperiods
under PAR of 180 µmol photons m^-2^ s^-1^ white LED, red or blue light.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^ growth rate trends
were similar to 250 µM O~2~. SS120 showed no growth under a 4 h
photoperiod for red spectral waveband, however under blue light, SS120
was able to grow (Fig \@ref(fig:SS120PARGAM)A). In contrast to the growth rate
trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^
experiments, SS120 grew under 4 and 8 h blue light and 16 h red light
photoperiods, however the growth rate decreased under 12 and 16 h white
LED light photoperiod treatments. Blue light treatments under PAR of 180
µmol photons m^-2^ s^-1^ showed growth only under an 8 h photoperiod.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16
h photoperiod and 180 µmol photons m^-2^ s^-1^ under white LED light
experiments were not performed due to time constraints.

SS120 did not reproducibly grow when sparged to the lowest O~2~ of 2.5
µM (Fig \@ref(fig:SS120PARGAM)A). 2.5 µM O~2~ growth rate experiments were not
conducted for 4 and 16 h photoperiods under peak PAR of 180 µmol photons
m^-2^ s^-1^ and for red light under peak PAR of 90 µmol photons
m^-2^ s^-1^, as no growth occurred when SS120 was exposed to 8 and 12 h
photoperiods.

@@

#  This is Fig10.tiff
NichesMap_cap <- glue("**Potential Future Niches for *Prochlorococcus* clades in a warming ocean.**",
                      "The maps shows current latitudinal distribution limits for *Prochlorococcus* (between green dashed lines), along with tropical and temperate  latitudinal regions (yellow bands).",
"Light levels, photoperiods and spectral bands, potentially permissive for growth of *Prochlorococcus* clades are determined by the interactions of latitude, season and depth attenuation of light.",
"**A.** Potential clade occupancies under Full Oxygen Zones (250 µM O~2~) are shown for 10-50 m and 100 m depths, in temperate vs. tropical regions, under depth-resolved photoperiods, approximating Winter Solstice, Equinox and Summer Solstice. Clade HLI is excluded from temperate depths by short photoperiods at Winter Solstice. Clades LLII/III and LLIV are excluded from shallow depths by excessive light.",
"**B.** Low Oxygen Zones (25 µM O~2~) drastically alter these potential occupancy patterns by excluding clade HLI from the temperate winter solstice at all depths, and by expanding the exploitation of higher light zones by Clade LLIV.",
"Potential clade occupancies for the southern temperate zone mirror the northern temperate zone occupancies.",
